---
title: "ggplot2 class code"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
pantheria<- read_csv("data/sub_PanTHERIA.csv") %>% 
  na.omit() %>% mutate(HabitatBreadth = as.character(HabitatBreadth))
```

## Basic syyntex for `ggplot2`

### `geom_histogram()`


```{r histogram, exercise=TRUE}
ggplot(pantheria,aes(x = log10(AdultBodyMass)))+
  geom_histogram()
```

### `geom_point()`

```{r point, exercise=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize)))+
  geom_point()
```

## Statistical Transformations

### example with geom_bar

```{r bar-error, exercise=TRUE, error=TRUE}
ggplot(pantheria,aes(Order,log10(AdultBodyMass)))+
  geom_bar()
```

### The fix

```{r bar, exercise=TRUE, error=TRUE}
ggplot(pantheria,aes(Order,log10(AdultBodyMass)))+
  geom_bar(stat = "identity")
```




## Scales

### Example with defult colors

Plotting the litter size as a function ob body mass grouped for activity cycle

```{r scales-basic, exercise=TRUE, error=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle))+
  geom_point()
```

### Plotting with `scale` option

```{r colorBrewer, exercise=TRUE, error=TRUE}
library(RColorBrewer)
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle))+
  geom_point()+
  scale_color_brewer(palette = "Dark2")
```


### Scaling without defining the parameter

```{r scale-not-defin, exercise=TRUE, error=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle))+
  geom_point()+
  scale_color_brewer(palette = "Dark2")+
  scale_shape()
```


The fix for this

```{r shape-scale-fix, exercise=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle,shape = ActivityCycle))+ #<<
  geom_point()+
  scale_color_brewer(palette = "Dark2")+
  scale_shape()
```


## Faceting

Plot the litter size as a function of body mass for each trophic level seperatly while keeping the shape-color scaling for acticity cycle.

`facet_wrap()`: define subsets as the levels of a single grouping variable

```{r facet_wrap, exercise=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle,shape = ActivityCycle))+
  geom_point()+
  scale_color_brewer(palette = "Dark2")+
  scale_shape()+
  facet_wrap(TrophicLevel~.)
```


`facet_grid()`: define subsets as the crossing of two grouping variables

```{r facet_grid, exercise=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle,shape = ActivityCycle))+
  geom_point()+
  scale_color_brewer(palette = "Dark2")+
  scale_shape()+
  facet_grid(TrophicLevel~HabitatBreadth)
```

Scale the axis in facet

```{r scale-axis, exercise=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle,shape = ActivityCycle))+
  geom_point()+
  scale_color_brewer(palette = "Dark2")+
  scale_shape()+
  facet_grid(TrophicLevel~HabitatBreadth,scales = "free") #<<
```


## Themes

```{r theme-light, exercise=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle,shape = ActivityCycle))+
  geom_point()+
  scale_color_brewer(palette = "Dark2")+
  scale_shape()+
  facet_grid(TrophicLevel~HabitatBreadth,scales = "free") +
  theme_light()#<<
```


```{r legend, exercise=TRUE}
ggplot(pantheria,aes(log10(AdultBodyMass),log10(LitterSize),
                     color = ActivityCycle,shape = ActivityCycle))+
  geom_point()+
  scale_color_brewer(palette = "Dark2")+
  scale_shape()+
  facet_grid(TrophicLevel~HabitatBreadth,scales = "free") +
  theme_light()+
  theme(legend.position = "none")#<<
```

## Position adjustments

**`position_dodge()`**

```{r position-dodge, exercise=TRUE}
ggplot(pantheria,aes(Order,log10(AdultBodyMass),fill = ActivityCycle))+
  geom_bar(stat = "identity",width = 0.6,
           position = position_dodge(width = 1))
```

**`position_jitter()`**

This is useful for **boxplot** and **violin** plot

```{r position_adjust,exercise = TRUE}
ggplot(pantheria,aes(ActivityCycle,log10(AdultBodyMass)))+
  geom_boxplot()+
  geom_jitter(alpha = 0.6,width = 0.2)
```

## Exercise


```{r, echo=FALSE}
sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30)
plotOutput("distPlot")
```

```{r, context="server"}
output$distPlot <- renderPlot({
  x <- faithful[, 2]  # Old Faithful Geyser data
  bins <- seq(min(x), max(x), length.out = input$bins + 1)
  hist(x, breaks = bins, col = 'darkgray', border = 'white')
})
```


